---
title: "The Impact of Walkability on Diabetes Prevelance in the Southern United States"
format:
  arxiv-pdf:
    keep-tex: true  
    linenumbers: false
    doublespacing: false
    runninghead: "A Preprint"
  arxiv-html: default
author:
  - name: Arkaprabho Bose
    affiliations:
      - name: Texas A & M University
        department: Undergraduate Program in Department of Computer Science
        address: 400 Bizzell Street
        city: College Station, TX
        country: USA
        postal-code: 77843
    orcid: 0000-0003-2402-304X
    email: abose0267@tamu.edu
    url: https://mm218.dev
  - name: Sebastian Oberg
    affiliations:
      - name: Texas A & M University
        department: Undergraduate Program in Department of Computer Science
        address: 400 Bizzell Street
        city: College Station, TX
        country: USA
        postal-code: 77843
  - name: Abhinav Cheruvu
    affiliations:
      - name: Texas A & M University
        department: Undergraduate Program in Department of Mathematics
        address: 400 Bizzell Street
        city: College Station, TX
        country: USA
        postal-code: 77843
abstract: |
  The diabetes epidemic in the United States, influenced by lifestyle elements such as walkability, presents a nuanced public health challenge. Our research shows that in southern regions, higher walkability indexes unexpectedly correlate with increased diabetes prevalence, which is contrary to the anticipated inverse relationship. This anomaly persists even when controlling for high blood pressure and low physical activity, which indicates significant regional variance. These findings present the necessity for tailored public health strategies and urban planning to effectively combat diabetes.
---

# Introduction {#sec-intro}

## Central Thesis Support
  

  
  What is so important about diabetes?
  
  Does a high walkability index always lead to a lower diabetes prevalence?
  
  In southern regions, a higher walkability index contributes to diabetes prevalence
  
  This is significant because it will show where diabetes needs more preventative measures
  
<!-- ## Paragraphs -->

  <!-- P1: This paragraph discusses the importance of diabetes, specifically the impact of it on a population and how common it is -->

  <!-- P2: Talk about how there are recent claims that lifestyle factors like living is a highly walkable diets, diets etc  can decrease diabetes cases. -->

  <!-- P3: State Thesis and Explain that it is important because it shows that the southern regions are not doing as well to prevent diabetes than other regions -->

<!-- ## Sentences -->




# Related Works

## Central Thesis Support



Exploring how location affects diabetes risk, focusing on two studies

One study shows that in Northeastern Germany, local factors like socio-economic status impact diabetes risk

Another study links diabetes with obesity and inactivity, stressing location-specific health solutions

These insights guide our analysis of walkability’s influence on diabetes in the Southern United States

<!-- Will reference these two from our journal entries: -->
<!--   - Do the risk factors for type 2 diabetes mellitus vary by location? A spatial analysis of health insurance claims in Northeastern Germany using kernel density estimation and geographically weighted regression -->
<!--   - Spatial Analysis of Incidence of Diagnosed Type 2 Diabetes Mellitus and Its Association With Obesity and Physical Inactivity -->

# Methods

## Central Thesis Support


In order to analyze diabetes as a response of walkability, we need good, clean data

In order to make sure that walkability is consistent as a coefficient, we choose multiple covariates that might be significant

We used GWR to create coefficient surfaces to analyze which parts of the country had the most impact on diabetes

We compared GWR with Random Forest to show that GWR is not overfitting, to prevent error


# Results

```{r echo={FALSE}}
source("./Functions/DataProcessing.R")

library(GWmodel)
library(dplyr)
library(ggplot2)
library(sf)
library(corrplot)
```

```{r echo={FALSE}}
loadedData <- loadData()
```

```{r echo={FALSE}}
cleaned_walkability = loadedData$original_walkability %>% clean_walkability()
cleaned_diabetes = loadedData$original_diabetes %>% clean_diabetes()
spatial_data = loadedData$spatial_data %>% clean_spatial()
new_dataset = merge(cleaned_walkability, cleaned_diabetes, by = "CountyFIPS", all.x = TRUE, all.y = TRUE) %>% subset( !(StateAbbr %in% c("AK", "HI")))
new_spatial_dataset = merge(new_dataset,spatial_data, by = "CountyFIPS", all.x = TRUE, all.y = TRUE)
```
```{r echo={FALSE}}
plot_correlation_matrix(new_dataset)
```

```{r echo={FALSE}}
new_spatial_dataset <- na.omit(new_spatial_dataset)
sf_spatial_data<- st_as_sf(new_spatial_dataset)
sf_spatial_data <- as(sf_spatial_data, "Spatial")
```

<!-- ```{r} -->
<!-- rf_pred <- runRandomForest(sf_spatial_data) -->
<!-- ``` -->

```{r echo={FALSE}}
ggplot() +
  geom_sf(data = new_spatial_dataset, aes(geometry = geometry,fill =DIABETES_CrudePrev), color="black",size=0.2) +
  scale_fill_viridis_c(name = "Diabetes Crude Prev") +
  labs(title = "Spatial Distribution of Diabetes Crude Prevalence")
```

```{r echo={FALSE}}
gwr_model_results = fit_gwr(sf_spatial_data)
model_df <- gwr_model_results$SDF %>% as("sf")
```

```{r echo={FALSE}}
  ggplot() +
    geom_sf(data = model_df, aes(geometry = geometry,fill = yhat), color="black",size=0.2) +
    scale_fill_viridis_c(name = "Diabetes Crude Prev") +
    labs(title = "Fitted Spatial Distribution of Diabetes Crude Prevalence")

  ggplot() +
    geom_sf(data = model_df, aes(geometry = geometry,fill = y), color="black",size=0.2) +
    scale_fill_viridis_c(name = "Diabetes Crude Prev") +
    labs(title = "Spatial Distribution of Diabetes Crude Prevalence")
```

```{r echo={FALSE}}
model_df %>%
 mutate("Predicted %Diabetes Prevalence" = yhat) %>%
  ggplot() +
  geom_sf(aes(fill = `Predicted %Diabetes Prevalence`),
          color = scales::alpha("black",
                                alpha = 0.1)) +
  scale_fill_gradientn(colours = terrain.colors(8),
                       limits = c(0, 25),
                       breaks = c(0, 10, 20)) +
  theme(text = element_text(size = 20),
        legend.position = "bottom") +
  labs(title = "Predicted Values for %Diabetes Prevalence by GWR w/ exponential kernel")
```

```{r echo={FALSE}}
model_df %>%
  mutate("%Diabetes Abs Err" = abs(yhat - y)) %>%
  ggplot() +
  geom_sf(aes(fill = `%Diabetes Abs Err`),
          color = scales::alpha("black",
                                alpha = 0.1)) +
  scale_fill_viridis_c(limits = c(0, 1),
                       breaks = c(0, 0.5, 1)) +
  theme(text = element_text(size = 20),
        legend.position = "bottom") +
  labs(title = "Absolute Prediction Error of %Diabetes by GWR w/ \nexponential kernel")
```

```{r echo={FALSE}}
model_df %>%
  mutate("NatWalkInd Coeff" = NatWalkInd) %>%
  ggplot() +
  geom_sf(aes(fill = `NatWalkInd Coeff`),color="black",size=0.2) +
 scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(title =
"Estimated Coefficient Surface of NatWalkInd")
```


```{r echo={FALSE}}
plots = data.frame(CountyFIPS = new_spatial_dataset$CountyFIPS,Walkability_Obs = new_spatial_dataset$NatWalkInd,HBP_Obs = new_spatial_dataset$BPHIGH_CrudePrev,walkability_coef= model_df$NatWalkInd, walk_impact = new_spatial_dataset$NatWalkInd * model_df$NatWalkInd, INTPTLAT=new_spatial_dataset$INTPTLAT, INTPTLON = new_spatial_dataset$INTPTLON, geometry =  new_spatial_dataset$geometry  ) %>%
  st_sf()

plots %>%
  ggplot() +
  geom_sf(aes(fill = walk_impact),color="black",size=0.2) +
 scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(title =
"Estimated Impact of NatWalkInd")

# example_long = model_df %>%
#   select(!c(Intercept), !contains("_TV"), !contains("_SE")) %>%
#   rename_with(~paste0(.x, "_coef")) %>%
#   bind_cols(new_spatial_dataset)
#   pivot_longer(data = NatWalkInd:CSMOKING_CrudePrev,
#                names_to = "Covariate",
#                values_to = "Coefficient")
```



```{r echo={FALSE}}
model_df %>%
  mutate("%HBP Prevalence Coeff" = BPHIGH_CrudePrev) %>%
  ggplot() +
  geom_sf(aes(fill = `%HBP Prevalence Coeff`),color="black",size=0.2) +
 scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  labs(title =
"Estimated Coefficient Surface of HBP")
```

```{r echo={FALSE}}
model_df %>%
  mutate("%Obesity Prevalence Coeff" = OBESITY_CrudePrev) %>%
  ggplot() +
  geom_sf(aes(fill = `%Obesity Prevalence Coeff`),color="black",size=0.2) +
 scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = mean(model_df$OBESITY_CrudePrev)) +
  labs(title =
"Estimated Coefficient Surface of Obesity")
```

```{r echo={FALSE}}
model_df %>%
  mutate("%Smoker Prevalence Coeff" = CSMOKING_CrudePrev) %>%
  ggplot() +
  geom_sf(aes(fill = `%Smoker Prevalence Coeff`),color="black",size=0.2) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = mean(model_df$CSMOKING_CrudePrev)) +
  labs(title =
"Estimated Coefficient Surface of Obesity")
```
# Discussion

## Central Thesis Support



Analyzing the relationship between walkability and diabetes prevalence in our study’s context

Comparing our findings with existing research to understand regional differences

Discussing the implications of our study for local health strategies

Highlighting the need for tailored interventions based on regional characteristics to combat diabetes

{{< pagebreak >}}

# References {.unnumbered}

