---
title: Spatial Analysis
subtitle: Sebastian Oberg, Arka Bose, Abhinav Cheruvu
format:
  homework-pdf: default
institute: Texas A\&M University
date: 2/27/2024
---

## Correlation Matrix
```{r}
# Load necessary libraries
library(dplyr)
library(corrplot) # correlation matrix visualization
library(ggcorrplot) # ggplot2 style visualization of correlation matrix

# Read datasets
data <- read.csv("../../Data/data.csv")
data2 <- read.csv("../../Data/500cities.csv")

# Prepare 'data' by padding and merging columns to create a unique identifier
data$STATEFP <- sprintf("%02d", as.numeric(data$STATEFP))
data$COUNTYFP <- sprintf("%03d", as.numeric(data$COUNTYFP))
data$TRACTCE <- sprintf("%06d", as.numeric(data$TRACTCE))
data$TractFIPS <- as.numeric(paste0(data$STATEFP, data$COUNTYFP, data$TRACTCE))

# Summarize 'data' to get mean of 'NatWalkInd' by 'TractFIPS'
averaged_data <- data %>%
  group_by(TractFIPS) %>%
  summarise(NatWalkInd = mean(NatWalkInd, na.rm = TRUE)) %>%
  ungroup()

# Select relevant columns from 'data2'
data2_relevant <- data2 %>% 
  select(TractFIPS, StateAbbr, DIABETES_CrudePrev, BPHIGH_CrudePrev, OBESITY_CrudePrev, LPA_CrudePrev, CSMOKING_CrudePrev)

# Merge 'averaged_data' with 'data2_relevant' on 'TractFIPS'
merged_data <- merge(averaged_data, data2_relevant, by = "TractFIPS", all.x = TRUE, all.y = TRUE)

# Filter for specific condition
merged_data <- merged_data %>% filter(StateAbbr == "CA")

# Calculate correlation matrix
cor_matrix <- cor(merged_data %>% select(-TractFIPS, -StateAbbr), use = "complete.obs")  # Handling missing values by using complete observations

# Output variable names (column names) used for correlation matrix
cat("Variables used for correlation matrix:\n")
print(names(merged_data %>% select(-TractFIPS, -StateAbbr)))

# Make correlation matrix
corrplot(cor_matrix, method = "circle", type = "upper", order = "hclust",
         tl.col = "black", tl.srt = 45,  # Text label color and rotation
         addCoef.col = "orange")  # Add correlation coefficients to plot
```

The correlation matrix above indicates strong positive correlations between obesity, physical activity, smoking, diabetes, and high blood pressure prevalence, with less correlation to the national walking index. Given these relationships, my team decided that a spatial autoregressive model (SAR) or a geographically weighted regression (GWR) could be appropriate to account for spatial dependencies and variations in the data.

<<<<<<< HEAD
## GWR Model
```{r}
install.packages("spgwr")
install.packages("spdep")
library(spgwr)
library(spdep)
library(sf)

spatial_data <- st_as_sf(merged_data, coords = c("longitude", "latitude"), crs = 4326) # placeholder

# Assuming 'DIABETES_CrudePrev' is dependent variable and rest are independent
formula_gwr <- DIABETES_CrudePrev ~ OBESITY_CrudePrev + LPA_CrudePrev + CSMOKING_CrudePrev + BPHIGH_CrudePrev

# Fit GWR model
gwr_model <- gwr(formula_gwr, data = merged_data, coords = cbind(spatial_data$lon, spatial_data$lat), bandwidth = 70, hatmatrix = TRUE)

# Print GWR model summary
print(summary(gwr_model))
```

## CAR Model
```{r}
# CAR models require a neighbors list
neighbors_list <- poly2nb(spatial_data) # placeholder

# Define formula for CAR model
formula_car <- DIABETES_CrudePrev ~ OBESITY_CrudePrev + LPA_CrudePrev + CSMOKING_CrudePrev + BPHIGH_CrudePrev

# Fit CAR model
car_model <- lagsarlm(formula_car, data = merged_data, listw = nb2listw(neighbors_list))

# Print CAR model summary
print(summary(car_model))
```
=======



```{r}
library(GWmodel)
library(sf)

merged_spatialdf <- as(merged_sf, "Spatial")

merged_gwr_bw <- bw.gwr(DIABETES_CrudePrev ~ OBESITY_CrudePrev + BPHIGH_CrudePrev + LPA_CrudePrev + CSMOKING_CrudePrev,
                        data = merged_spatialdf,
                        kernel = "exponential",
                        parallel.method = "omp")

merged_gwr <- gwr.basic(DIABETES_CrudePrev ~ OBESITY_CrudePrev + BPHIGH_CrudePrev + LPA_CrudePrev + CSMOKING_CrudePrev,
                        data = merged_spatialdf,
                        bw = merged_gwr_bw,
                        kernel = "exponential",
                        parallel.method = "omp")

gwr_results_sf <- merged_gwr$SDF %>% as("sf")


```

>>>>>>> 85aa983 (Added GWR Model)
